A Summery about reasoning with logic

### 2015-10-31 reasoning/summery.tex
added a section about jargon and deleted the proof ad absurdom started explaining induction with an almost trivial example.

@@

So there are 2 kinds of proving. The first one is writing a proof in the system
this should be rather easy, although you have to be carefull to follow the
rules to the letter. I'm going to ignore this one since its just using the
rules, which you've done tons of times as a programmer, so it shouldn't be
that strange.

The second one is the more troublsome, its proving the system itself. This
is also split up into 3 different aspects. They all have in commen that
it should be done in natural langauge. But, there is a structure to it
I'm going to try and give a complete overview of this structure.

First you have soundness, then there is completness, and finally you can be
asked to proof certain properties of a system.

@@

\section{proving soundness}
This basically means your system doesn't modify the value you put in.
If you put in truth, apply a bunch of rules from your system, then you should
get out truth.

\subsection{Structure of your proof}
First write down your overal strategy of the proof \autoref{strategies}. For
soundness this is often `proof by induction' \autoref{induction}.

Then write down the lemmas you're gonna use. A lemma is a little proof, often
derived from the properties of the system you're proving. You could see it
as a basic argument. When proving soundness you want to connect this to the
semantics, for propositional systems the semantics are truth tables, for kripke
models it are kripke frames. The semantics are the thing your talking about
in your logical system.
%I can see how I dind't understand all this crap the first time I saw it,
% its freaking packed with definitions, and even if you understand all the
% definiontions, its still hard
Make sure that you've proven the soundness of the axioms *and* of the rules
with help of the semantics.

Finally ty up the lemmas together in the conclusion, this is done with help
of the proofing strategy.

Finish with a sentence like: `This had to be shown'.

@@

\subsection{Induction}
\label{induction}

\section{notations}
\label{notations}
There is no standart in notating things. Every logician seems to use a
slightly different notations. Keep this in mind. But I think the connectives
are mostly the same. You probably need to memorize these.

>>>>

\newcommand{\brcell}[2][l]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

@@

Proving is in meta space. Its a little like programming in that you have to
be specific. There are lots of ways to prove things. In this report I will
discuss a select few to give an idea about what is going on. But first
I'll start covering with some basics.

\subsection{Jargon}
In here I will discuss some jargon logicians use. It may seem silly but knowing
the precise defintion of these words makes understanding proofs a lot easier.

\begin{tabular}{ll}
	Word & definition \\ \toprule
	Formula & The letters, but also letters combined with conectives (its
	recursive) \\
	Axiom & Something which is undoubtibly valid. proofs work from a
	complex formula towards axioms. \\
	Theorom & the thing you're trying to proof. \\
	Lemma & \brcell{a hypothosis which you're trying to prove, you usually
		define \\
	a lemma to reference to later. Its like a little theorom} \\
	Sequent & A thing with a sequent connective (in my case either
	$\vdash$ or $\Rightarrow$) \\
	Derivable & if X is derivable in system A, then you can proof x with
	system A
\end{tabular}

@@

\section{Structure of proofs}
\label{structure of proofs}
Now of course different poeple have different kinds of notations. The logicians
don't really seem to adhere to standarts. But that's fine, you just have to
get used to conflicting symbols.

\subsection{Find references}
Before you do anything check if you can't refer to any existing proofs.
This saves a lot of time. You can just ignore writing a proof if you can
write down a reference. (I bet this is allowed on the exam as long you
can refer to the stuff you wrote down yourself).

\subsection{Head of a proof}
In the head of the proof you don't do any work, you just state what we're
talking about and what you're going to do. So you start with the definitions.
Then you say what you're going to proof (by using the defintions), so here
you just repeat the question. And finally you mention your strategy.

\subsection{Body of the proof}
In here you work out the proof with help of the strategy.
See the \autoref{strategies} section.

\subsection{End of the proof}
Just write down ` QED' or ` this is what had to be shown' even if you're not
certain what you're doing.

@@

\subsection{Induction}
\label{induction}
\subsubsection{When to use it?}
You can use induction if you have to proof something for countibly infinite
things, but there exists a base method that can validate if its true.
So basically all thigns that can and make sense be resolved with a
recursive function.

\subsubsection{How to use it?}
In here you try to prove something by starting with a single case n.
For all n (*), n=1, supose (*) for n, then show (*) for n+1.

Fat chance you'll understand it based on that. Its way to abstract.
But I'll show how to use it with help of several examples. \emph{cursive
text will be comments}

\paragraph{Proof that formulas always have an even number of brackets.}
\emph{So space cowboy, we have to write a proof, before panicking remember
	the boiler plate in \autoref{structure of proofs}, we may get some
points for doing this, its at least polite}

\subparagraph{Definitions}
A formula can be a propositional variable, in which case it has zero brackets.
A formula can also be multpile proporsitional variables, connected with connectives,
in this case it has the connective count * 2 brackets, because for every connective
2 brackets are added.
\subparagraph{Strategy} Every formula has a even number of brackets, this will
be proven with help of iniduction.
\subparagraph{Proof}

\subsection{Tabloo or truth tree}
\subsubsection{When to use it?}
The tabloo method comes to mind if you have to proof things about systems
that use trees, and where the values of the leaves of the tree trincle down
to the root of the tree.

\subsubsection{How to use it?}
In this strategy you create a tree, starting from the axioms which is either
true or false, you go down step by step and say what happens. In most systems
if one of the axioms is false it means the conclusion is false to.

\section{Notations}
\label{notations}
There is no standart in notating things. Every logician seems to use a
slightly different notations. Keep this in mind. But I think the connectives
are mostly the same. You probably need to memorize these.
